---
title: "Project"
output:
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(forecast)
library(ggplot2)
library(urca)
```

## Data & Descriptive Plots
```{r}
#Import Data - Japan Real GDP, Quarterly Data, seasonaly adjusted
gdp <- read.csv('JPNRGDPEXP.csv')
#Turn the data into time series format
gdp.ts <- ts(gdp$JPNRGDPEXP,start=c(1994,1),freq=4)
```


```{r}
#Simple Data Plot
plot(gdp.ts,  ylab = "Real GDP", xlab = "Year", bty = "l",xaxt="n")
axis(1, at = seq(1994, 2020, 1)) 
grid()
```


```{r}
#Log Difference Plot
gdp.log <- log(gdp.ts)
gdp.ts.diff <- diff(gdp.log, lag=1)
plot(gdp.ts.diff,  ylab = "GDP growth rate", xlab = "Year", bty = "l",xaxt="n")
axis(1, at = seq(1994, 2020, 1)) 
grid()
```
From the simple time plot above, we can see that Japan GDP is following an upward slop trend with a gap in 2009. After we convert data to GDP growth rate(log difference), there's no clear trend in it. Sometimes GDP growth rates fall above 0, which means there was a positive GDP growth. However, sometimes GDP growth rate falls below 0 which indicates decreasing GDP. For instance, the most remarkable drop of growth rate for Japan GDP is over 4% drop in 2009. There seems to be more volatility in GDP after 2008 comparing with GDP before 2008.

## Preparation

```{r}
#Look into ACF and PACF plots in order to gain better understannding of the data
acf(gdp.ts)
pacf(gdp.ts)
```
Plots of ACF and PACF above can help with determining suitable model. In this case, because ACF shows autocorrelation dies out slowly as lag increases, while PACF drops to 0 suddenly after lag 1, it could be appropriate to use AR(1) model here. 



```{r}
# DF Test for stationarity
# Use trend ADF test because there's a clear trend in the GDP data, "BIC" since the dataset is large
df.test <- ur.df(gdp.log,type="trend",selectlags="BIC")
print(summary(df.test))
sumStats <- summary(df.test)
#Third statistic since type = 'trend'
teststat <- sumStats@teststat[3]
# critical value at 5 percent
critical <- sumStats@cval[2]
teststat > critical

```
Cannot reject the null, so this is non stationary. It is a Random walk plus drift, which make sense given that there is a growth rate. Thus, I decide to use Arima model since Arma can only be used when the time series data is stationary. With non stationary data, Arima can be used to take the first difference and forecast based on the difference. 

## Model Selection

```{r}
#Use data until 2012 as training set and the rest as validation set
gdp.train.ts <- window( gdp.log, start=c(1994,1),end=c(2012,4))
gdp.valid.ts <- window( gdp.log, start=c(2013,1))
# length of full data set
T <- length(gdp.log)
# length of training data set
T2 <- length(gdp.train.ts)

# Arima model with trend 
mod1 <- Arima(gdp.train.ts,order=c(1,1,0),include.constant = TRUE)
mod1.fcast <- forecast(mod1,h=T-T2,level=95)

#plot forecase
plot(mod1.fcast, xlab="Year",ylab="Log(GDP)")
lines(gdp.log)
lines(gdp.log,lwd=1)
legend(x="topleft",c("Data","Forecast"),col=c("black","blue"),lwd=2)
grid()
```


```{r}
#residual diagnosis: uncorrelated
plot(mod1$residuals)
ggAcf(mod1$residuals)
```



```{r}
# ARIMA(d=1) model, because 1994-2020 is a lonng time horizon, use bic here
mod2 <- auto.arima(gdp.train.ts,d=1,ic="bic",seasonal=FALSE)
mod2.fcast <- forecast(mod2,h=T-T2,level=95)
plot(mod2.fcast,xlab="Year",ylab="Log(GDP)")
lines(gdp.log)
legend(x="topleft",c("Data","Forecast"),col=c("black","blue"),lwd=2)
grid()
```
The model that auto.arima generated is Arima(0,1,0), which is just a random walk forecasting and can be used as a baseline model to compare with my arima model with drift.


```{r}
# Following the suggestion in peer review, add another double exponential filtering/ Holt model wtih trend as comparison with my arima model as well
ses <- ets(gdp.train.ts, model="AAN")
ses.pred <- forecast(ses, h=T-T2, level=95)
plot(ses.pred,xlab="Year",ylab="Log(GDP)")
lines(gdp.log)
legend(x="topleft",c("Data","Forecast"),col=c("black","blue"),lwd=2)
grid()
```

```{r}
# Compare the accuracy of three models
accuracy(mod1.fcast,gdp.log)
accuracy(mod2.fcast,gdp.log)
accuracy(ses.pred,gdp.log)
# Conduct DM test to see if my model is a better model comparing with baseline Arima(0,1,0) & the Holt model
dm.test(residuals(mod1.fcast),residuals(ses.pred),alternative="less")
dm.test(residuals(mod1.fcast),residuals(mod2.fcast),alternative="less")

```
By comparing the accuracy of three models and conductinng DM tests, I found that although my Arima(1,1,0) with drift model is not significantly better than the other two models, its RMSE performance is better than the Holt model and way better than the Arima(0,1,0) model. The same conclusion can also be drawn from the DM tests since p-value of the DM test between Arima(1,1,0) with drift and Arima(0,1,0) is much lower than the p-value of the DM test between Arima(1,1,0) with drift and Holt model, indicating that it's more likely that Arima(1,1,0) with drift is a better model than Arima(0,1,0).
